---
title: "Data Comparison"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This report compares the current RRDS flow calculations to flows
calculated from the culvert head loss equations provided by the
`culvrt` package.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
library(lubridate)
]```

First, load system data.

```{r}
dfile = system.file("sample-data/example.csv", package = "culvrt")

raw.data = read_csv(dfile, na = c("", "(null)"), guess_max = 1e4)

cleaned.data = raw.data %>%
  mutate(head.loss = screen - pond)
```

Now calculate the velocity in each culvert based on the gate height:

```{r}
velocity.data = cleaned.data %>%
  select(-velocity.plc) %>%
  pivot_longer(matches("GATE"), 
    names_to = "gate",  values_to = "height") %>%
  na.omit() %>%
  filter(head.loss > 0, height > 0) %>%
  mutate(
    velocity.eq1996 = map2_dbl(head.loss, height,
      ~ uniroot(root_velocity_hist, c(0, 50), head.loss = .x, h = .y,
          n = 0.026, ks.table = brater_ks())$root),
    velocity.eq2011 = map2_dbl(head.loss, height,
      ~ uniroot(root_velocity, c(0, 50), head.loss = .x, h = .y,
          n = 0.022, ks.table = usda_ks())$root),
    velocity.eq2021 = map2_dbl(head.loss, height,
      ~ uniroot(root_velocity_alt, c(0, 50), head.loss = .x, h = .y,
          n = 0.022, ks.table = usda_ks())$root)
  )
```

Convert velocity to flow and then sum across gates:

```{r}
flow.data = velocity.data %>%
  mutate(across(starts_with("velocity"), velocity_to_flow,
    .names = "{gsub('velocity', 'flow', .col)}")) %>%
  group_by(time) %>%
  summarize(across(starts_with("flow"), sum), .groups = "drop")
```

Join the flow data back to the system data and calculate the fish
screen velocity:

```{r}
complete.data = cleaned.data %>%
  left_join(flow.data, by = "time") %>%
  select(-starts_with("GATE")) %>%
  mutate(across(starts_with("flow"), ~ .x / (120 * (tide + 7.2)),
    .names = "{gsub('flow', 'velocity', .col)}")) %>%
  select(-starts_with("flow")) %>%
  pivot_longer(starts_with("velocity"), names_to = "method",
    values_to = "velocity") %>%
  mutate(
    velocity = if_else(is.na(velocity), 0, velocity),
    method = if_else(str_detect(method, "plc$"), "PLC",
      paste(str_extract(method, "[0-9]{4}"), "equation"))
  )

complete.flow.data = cleaned.data %>%
  left_join(flow.data, by = "time") %>%
  select(-starts_with("GATE")) %>%
  pivot_longer(starts_with("flow"), names_to = "method",
    values_to = "flow") %>%
  mutate(
    flow = if_else(is.na(flow), 0, flow),
    method = if_else(str_detect(method, "plc$"), "PLC",
      paste(str_extract(method, "[0-9]{4}"), "equation"))
  )
```

```{r}
#save(complete.data, file = "data-compare.RData")
```


```{r}
complete.data %>%
  filter(between(time, as_datetime("2019-08-30 00:00:00"),
    as_datetime("2019-10-29 15:00:00"))) %>%
  ggplot() +
    aes(x = time, y = velocity, color = method) +
    geom_line(size = 0.5, alpha = 0.5) +
    scale_color_brewer(NULL, palette = "Set1") +
    scale_x_datetime(NULL, date_labels = "%H:%M\n%Y-%m-%d",
      expand = expansion(0.01)) +
    scale_y_continuous("Fish screen velocity (fps)") +
    theme_bw() +
    theme(
      legend.position = c(0.15, 0.85),
      legend.background = element_rect(color = "black")
    )

plotly::ggplotly(dynamicTicks = TRUE)

# subsets

sub1 = complete.data %>%
  filter(between(time, as_datetime("2019-09-21 06:00:00"),
    as_datetime("2019-09-22"))) %>%
  mutate(group = 1L)

sub2 = complete.data %>%
  filter(between(time, as_datetime("2019-10-05 06:00:00"),
    as_datetime("2019-10-06 02:00:00"))) %>%
  mutate(group = 2L)

sub3 = complete.data %>%
  filter(between(time, as_datetime("2019-10-28 13:00:00"),
    as_datetime("2019-10-29 15:00:00"))) %>%
  mutate(group = 3L)


bind_rows(sub1, sub2, sub3) %>%
  ggplot() +
    aes(x = time, y = velocity, color = method) +
    geom_line(size = 1) +
    scale_color_brewer(NULL, palette = "Set1") +
    scale_x_datetime(NULL, date_labels = "%H:%M\n%Y-%m-%d") +
    scale_y_continuous("Fish screen velocity (fps)") +
    theme_bw() +
    facet_wrap(~ group, ncol = 1, scales = "free_x") +
    theme(
      legend.position = c(0.15, 0.9),
      legend.background = element_rect(color = "black"),
      strip.background = element_blank(),
      strip.text.x = element_blank()
    )

# new sub


complete.data %>%
  filter(between(time, as_datetime("2019-10-10 09:00:00"),
    as_datetime("2019-10-11 21:00:00"))) %>%
  ggplot() +
    aes(x = time, y = velocity, color = method) +
    geom_line(size = 1, alpha = 0.8) +
    scale_color_brewer(NULL, palette = "Set1") +
    scale_x_datetime(NULL, date_labels = "%H:%M\n%Y-%m-%d") +
    scale_y_continuous("Fish screen velocity (fps)", limits = c(0, 1.05)) +
    theme_bw() +
    theme(
      legend.position = c(0.15, 0.85),
      legend.background = element_rect(color = "black")
    )
```
